The Increased Effect of Elections and Prime Ministers on Discussion in the Australian Federal Parliament between 1901 and 2018

### 2019-04-01 outputs/docs/paper/paper.Rmd
Update paper

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Effect of Elections and Prime Ministers on Discussion in the Australian Federal Parliament (1901--2018)"
thanks: "Thank you to Chris Cochrane, Dan Simpson, Jill Sheppard, John McAndrews, John Tang, Leslie Root, Martine Mariotti, Matt Jacob, Matthew Kerby, Myles Clark, Ruth Howlett, Tianyi Wang, Tim Hatton, and Zach Ward for their helpful suggestions; and to the UC Berkeley Demography Department for the use of their computing resources. We are grateful for the many excellent comments that we received from seminar participants at the ANU SPIR, the ANU RSE, the Australian Parliamentary Library, the Max Planck Institute for Demographic Research, the UofT Political Behavior Group, and the 2019 PSA Political Methodology Conference. Comments and suggestions on the `r format(Sys.time(), '%d %B %Y')` version of this paper welcome at: rohan.alexander@anu.edu.au."
author:
- name: Rohan Alexander
  affiliation: Australian National University
- name: Monica Alexander
  affiliation: University of Toronto
abstract: "A large part of politics consists of reacting to events, but not all events provoke a reaction. As such, understanding political behaviour requires examining the reaction to events and how this changes over time. We systematically analyse how discussion in the Australian Federal Parliament changes in response to two types of events: elections and changed prime ministers. To do this we first create a new dataset of what was said in the Australian Federal Parliament from 1901 through to 2018 based on available public records. We reduce the dimensionality of this dataset using a correlated topic model, relate those topics to the Comparative Agendas Project, and then analyse the effect of these two types of events using a Bayesian hierarchical Dirichlet model that has several advantages over existing models. We find that: changes in prime minister tend to be associated with topic changes even when the party in power does not change; and elections that do not result in a change in prime minister are rarely associated with topic changes in the first half of our sample, but have been increasingly significant since the 1980s."
keywords: "text-as-data, Australian politics, unsupervised machine learning, Bayesian hierarchical Dirichlet model, Comparative Agendas Project"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no

@@

names(key_events_results_cols) <- c("Dates", "", "", "", "", "", "", "")

@@

We systematically analyse the text in two stages, each of which takes advantage of a different statistical technique. First, to reduce the dimensionality of our dataset we use a topic model that allows correlation between topics. Second, we analyse those topics using a Bayesian hierarchical Dirichlet model to examine changes at various types of events. We explicitly consider two types of events: changes in prime minister and elections. After taking these effects into consideration, we then explore other significant events.

We find that: firstly, changes in prime minister tend be associated with topic changes even when the party in power does not change. For instance, the change from Hughes to Bruce in 1922, Menzies to Holt in 1966; Hawke to Keating in 1991; and Rudd to Gillard in 2010 are all associated with significant changes in topics despite no change in the party in power. Secondly, as expected, elections where the party in power also changes, such as Fisher in 1910 and 1914, Menzies in 1949, Whitlam in 1972, Fraser in 1975, Hawke in 1983, Howard in 1996, Rudd in 2007, and Abbott in 2013 are associated with topic changes, but the 1974 Whitlam, 1980 Fraser, and the 1998 and 2004 Howard re-elections, stand out as elections where the prime minister did not change but there was a significant change in topics.

@@

[^XMLfootnote]: Tim Sherratt makes Commonwealth XML records for 1901 to 1980 available as a single download at: http://historichansard.net/. Commonwealth XML records from 1998 to 2014 are available from Andrew Turpin’s website, and from 2006 through to today from Open Australia’s website. The records can also be downloaded from the Australian Hansard website or the website can be scraped. We do not use the XML records or scrape the website because those records are known to be incomplete but the extent of how incomplete they are is unknown. The trade-off for a more-complete record is the errors introduced by having to parse the PDFs.

We use the official PDF release and the formatting of the Hansard records changes over time. We use scripts written in R [@R2018] to convert the PDFs into daily text records.[^scriptsfootnote] Some error is introduced at this stage because many of the records are in a two-column format that need to be separated, and the PDF parsing is not always accurate, especially for older records. An example of the latter issue is that 'the' is often parsed as 'thc'. These errors are corrected when they can be identified.

[^scriptsfootnote]: The exact scripts that we use are available on request or via the GitHub repository for this paper. The scripts are primarily based on: the \texttt{PDFtools} R package of @Ooms2018pdftools; the \texttt{tidyverse} R package of @Wickham2017; the \texttt{tm} R package of @FeinererHornik2018; the \texttt{lubridate} R package of @GrolemundWickham2011; the \texttt{tidytext} R package of @SilgeRobinson2016; and the \texttt{stringi} R package of @Gagolewski2018. The functions of those packages are augmented by: the \texttt{furrr} R package of @VaughanDancho2018; and the \texttt{tictoc} R package of @Izrailev2014. The \texttt{hunspell} R package of @Ooms2017 is used to help find spelling issues; and the \texttt{quanteda} R package of @Benoit2018 is used to compound multiword expressions.

@@

[^datasourcefootnote]: Our dataset is available at: https://github.com/RohanAlexander/hansardr. While we are making our data public in an attempt to help other researchers, we cleaned the dataset toward the requirements of this paper. If you have different requirements or have suggestions for how we can improve the dataset for other uses please contact us.

@@

We first use a Correlated Topic Model (CTM) [@BleiLafferty2007] to obtain estimated topic distributions over time. We consider these topic distributions as inputs that can be analysed by another model. Thus, the second modelling step involves using a Bayesian hierarchical Dirichlet model to analyse changes in the topic distributions (obtained from the first step) in relation to events of interest.

@@

Other work such as @BaumgartnerJones1993 and @DowdingHindmoor2010 addressed this problem by creating a standardised codebook of policy categories and sub-categories and then manually assigning text to topics as appropriate. This approach ensures the categorisation is reasonable but as it is a manual process the size of the text that can be categorised is limited. @AshMorelliOsnabru2018 combined the best of both in the context of the New Zealand Hansard, using a supervised machine learning approach that would likly be beneficial in the Australian Hansard case as well.

In order to effectively categorise the topics of the entire Hansard, we use topic modelling, a statistical technique which aims to extract the underlying or latent 'topics' from a collection of texts. In particular, we use a method which is similar to Latent Dirichlet Allocation (LDA), first developed by @Blei2003latent. One advantage of using the Dirichlet distribution is that it formalises the idea that there is a trade-off in what is talked about in parliament given the limited time. A multinomial approach would not have this property.

@@

As such, we use the Correlated Topic Model (CTM) of @BleiLafferty2007 to obtain topic distributions. The CTM is a modification of LDA that allows for correlations between topics. More detail about the CTM is in Appendix \@ref(correlatedtopicmodelsection).

@@

The CTM output of interest is the proportion of each topic appearing in each document. The aim of this stage of the modelling process is to analyse how the distribution of topics changes in relation to different types of events. But with many topics for each of the roughly 14,680 chamber-sitting-days the data are still too noisy to easily visualise changes around events.

@@

Define $\theta_{cdp}$ to be the proportion of topic $p$ on day $d$ in chamber $c$ (where chamber refers to either the House of Representatives or the Senate). Note that the $\theta_{c,d,1:P}$ for $p = 1,2,\dots, P = 80$ are equal to the estimated values of $\theta_{cd}$ from the CTM. We assume that the majority of variation in topics by day is across sitting periods, $s$, where a sitting period is defined as any group of days that are less than one week apart. Using this definition, there are 822 sitting periods over the period 1901 to 2018 inclusive. Appendix \@ref(hansardsummarystatistics) contains more information about the sitting patterns over the course of the year, which have changed considerably since Federation.

@@

The goal of the model is to relate these proportions to the current prime minister $g$ and election period $e$, assuming that an 'election effect' would be at its peak in the days close to the election, then decay over the period. The mean parameters $\mu_{c,s,p}$ are modelled on the log scale as:

$$
\log \mu_{c,s,p} = \alpha_{g[c,s],p} + \beta_{e[c,s],p} \cdot (N_{s[e]} - s) + \delta_{c,s,p}
$$
where: $\alpha_{g[c,s],p}$ is the mean effect for prime ministers $g$ (which covers sitting period $s$ in chamber $c$) and topic $p$; $\beta_{e[c,s],d,p}$ is the effect of election $e$ (which occurs in sitting period $s$, in chamber $c$) for topic $p$; $N_s$ is the total number of sitting periods in election period $e$; and $\delta_{c,s,p}$ is a random, or levels, effect for each chamber, sitting period and topic.

@@

Firstly, we describe the results of the CTM approach, which defined 80 topics over the period 1901 to 2018. We then describe the results of the Bayesian analysis model, which identified prime ministers, elections and other events that were associated with a change in the topics discussed.

@@

LDA output defines a topic as a distribution of probabilities over words. In Table \@ref(tab:topwords) we display the ten words with the highest association for each of the 80 topics, but the topics are defined by probability distributions over all words. LDA does not apply labels to each topic, or collection of words, instead this has to be done by inspection. The topics cover areas such as budgets, demography, transport and infrastructure, war and conflict, health, education, agriculture, and trade. Similar to when topic models are run using the parliamentary text records of other countries, there are also some topics that seem to be about procedural or day-to-day matters, such as Topics 7 or 9. As expected, some topics seem to somewhat overlap with their content: for instance, Topics 4, 26, 28, 30 and 66 all relate to war and conflict.

@@

In the final column of Table \@ref(tab:topwords) we summarise the results of our 80-topic model using the categories of the Comparative Agendas Project (CAP), as summarised by @Bevan2017.\footnote{We use Version 1.0 of the Master Codebook, as at 31 July 2014 available at \url{http://sbevan.com/cap-master-codebook.html}.} We do this firstly, to assist other researchers in understanding our results, by trying to relate them to a a more well-known approach. We also do this because it allows us to further reduce our 80-topic model to only 21 topics.\footnote{Although the CAP Codebook topic numbering goes to 23, there are only 21 topics because there is no topic 11 or topic 22. Our 80 topics did not relate to CAP topics 21 or 23, meaning that we have 19 CAP-based categories.} It is these 19 CAP groupings to which we apply our analysis model.

@@

We are interested in considering the effect of various political and other events on what is talked about in the Australian Federal Parliament. As discussed in Section \@ref(analysismodelexplanation), the modelling process takes the topic proportions estimated by the CTM, and examines the association between these topic distributions and outside events. That is, the topic model outputs are inputs for our analysis model.

There are several outputs of interest from this modelling stage. For example, the model provides estimates of topic prevalence by each sitting period. This nicely illustrates how the topics change over time, as the daily estimates tend to be quite variable, but using periods defined by prime ministers or elections tend not to provide enough variability. For instance, examining Topics 4, 26, 28, 30 and 66, which have to do with war and conflict illustrates Australia's involvement in World War I, World War II, the Korean War, the Vietnam War, the First Gulf War, the War in Afghanistan and the Second Gulf War (Figure \@ref(fig:wartopicsgraph)).

```{r wartopicsgraph, cache = TRUE, echo = FALSE, fig.cap = "Model estimates of topic prevalance by sitting period for those related to war and conflict", out.width = '100%'}
knitr::include_graphics("../../figures/war_stack.pdf")
```

@@

```{r mugov, cache = TRUE, echo = FALSE, fig.cap = "Level effects for prime ministers by selected topics.", out.width = '100%'}
knitr::include_graphics("../../figures/mu_gov.pdf")
```

```{r muelection, cache = TRUE, echo = FALSE, fig.cap = "Level effects for election periods by selected topics", out.width = '100%'}
knitr::include_graphics("../../figures/mu_election.pdf")
```

@@

<!-- Table \@ref(tab:keyeventsresultstable) summarises the days where parliamentary discussion was significantly different from the rest prevailing in that week. -->

<!-- ```{r keyeventsresultstable, echo = FALSE, results = 'asis'} -->

<!-- key_events_results_cols %>% -->
<!--   kable(booktabs = T, caption = "Days that were significantly different to other days in their sitting period") %>% -->
<!--   kable_styling(font_size = 10) %>%  -->
<!--   footnote( -->
<!--     general = "These days were significantly different to others in their sitting period after taking election and prime minister effects into consideration.",  -->
<!--     footnote_as_chunk = T,  -->
<!--     threeparttable = TRUE) -->

@@

Of the 36 prime ministers over this period, we find that 20 of them are significantly different to the prime minister that preceded them. However, four of these results -- the significance of the Page term, the first Menzies term, the Chifley term and the Abbott term -- are likely due to the short length of either the prime minister or the predecessor and should be ignored, leaving only 16. The earliest of these are the first Fisher term, the Cook term and the third Fisher term, which were different to the second Deakin term, the second Fisher term, and the Cook term, respectively. To a certain extent this is likely due to changes due to World War I.

@@

Menzies was succeeded  Holt in January 1966. This is an example where there was a change in prime minister without an election, as the next election only happened in November 1966. We find that Holt is different to Menzies. In Figures \@ref(fig:muelectionall) we compare the topics during Menzies' final term with the topics of Holt.

```{r muelectionall, cache = TRUE, echo = FALSE, fig.cap = "Differences between various elections", out.width = '100%'}
knitr::include_graphics("../../figures/mu_election-differences.pdf")
```

@@

One advantage of our analysis model compared with using the STM approach is that we can create a measure that is equivalent to testing for outliers in a model where the underlying variables were not latent. The results of this reduction in supervision are promising, but suggest the specifics of our process may need further refinement. For instance, our approach appropriately identifies the sitting day that first follows 11 September 2001 and 12 October 2002, which were the dates of the 9/11 Attacks and Bali Bombings respectively. But there are many dates that we would have expected to be identified that were not, and similarly some of the dates that were identified are surprising. When we examine this we find that some of them, such as 4 February 2009, are associated with significant legislation. However the over-weighting of dates in the first half of our sample in Table \@ref(tab:keyeventsresultstable) highlights further work is needed to improve this approach. For instance, our approach may not be appropriately considering step-changes or it may be that our identification of sitting periods is not appropriate for the entire period.

>>>>

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Increased Effect of Elections and Prime Ministers on Discussion in the Australian Federal Parliament between 1901 and 2018"
thanks: "Thank you to Chris Cochrane, Dan Simpson, Jill Sheppard, John McAndrews, John Tang, Leslie Root, Martine Mariotti, Matt Jacob, Matthew Kerby, Myles Clark, Ruth Howlett, Tianyi Wang, Tim Hatton, and Zach Ward for their helpful suggestions; and to the UC Berkeley Demography Department for the use of their computing resources. We are grateful for the many excellent comments that we received from seminar participants at the ANU SPIR, the ANU RSE, the Australian Parliamentary Library, the Max Planck Institute for Demographic Research, the U of T Political Behavior Group, and the 2019 PSA Political Methodology Conference. The code and data for this paper is available at: https://github.com/RohanAlexander/hansard. Comments and suggestions on the `r format(Sys.time(), '%d %B %Y')` version of this paper welcome at: rohan.alexander@anu.edu.au."
author:
- name: Rohan Alexander
  affiliation: Australian National University
- name: Monica Alexander
  affiliation: University of Toronto
abstract: "A large part of politics consists of reacting to events, but not all events provoke a reaction. As such, understanding political behaviour requires examining the reaction to events and how this changes over time. We systematically analyse how discussion in the Australian Federal Parliament changes in response to two types of events: elections and changed prime ministers. To do this we first create a new dataset of what was said in the Australian Federal Parliament from 1901 through to 2018 based on available public records. We reduce the dimensionality of this dataset using a correlated topic model, relate those topics to the Comparative Agendas Project, and then analyse the effect of these two types of events using a Bayesian hierarchical Dirichlet model that has several advantages over existing models. We find that: changes in prime minister tend to be associated with topic changes even when the party in power does not change; and elections that do not result in a change in prime minister are rarely associated with topic changes in the first half of our sample, but have been increasingly significant since the 1980s."
keywords: "text-as-data, Australian politics, unsupervised machine learning, Bayesian hierarchical Dirichlet model, Comparative Agendas Project"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no

@@

names(key_events_results_cols) <- c("Dates", "", "", "", "", "")

@@

We systematically analyse the text in two stages, each of which takes advantage of a different statistical technique. First, to reduce the dimensionality of our dataset we use a topic model that allows correlation between topics, and then further reduce those topics to match the Comparative Agendas Project. Second, we analyse those topics using a Bayesian hierarchical Dirichlet model to examine changes at various types of events. We explicitly consider two types of events: changes in prime minister and elections.

We find that: firstly, changes in prime minister tend be associated with topic changes even when the party in power does not change. For instance, the change from Hughes to Bruce in 1923, Forde to Chifley in 1945; Menzies to Holt in 1966; Hawke to Keating in 1991; and Rudd to Gillard in 2010 are all associated with significant changes in topics despite no change in the party in power. Secondly, as expected, elections where the party in power also changes, such as Fisher in 1910 and 1914, Menzies in 1949, Hawke in 1983, and Howard in 1996 are associated with topic changes, but the 1974 Whitlam, and 1984 Hawke re-elections stand out as elections where the prime minister did not change but there was a significant change in topics.

@@

[^XMLfootnote]: Tim Sherratt makes Commonwealth XML records for 1901 to 1980 available as a single download at: http://historichansard.net/. Commonwealth XML records from 1998 to 2014 are available from Andrew Turpin’s website, and from 2006 through to today from Open Australia’s website. The records can also be downloaded from the Australian Hansard website or the website can be scraped. We do not use the XML records or scrape the website for this paper because those records are known to be incomplete but the extent of how incomplete they are is unknown. The trade-off for a more-complete record is the errors introduced by having to parse the PDFs.

We use the official PDF release and the formatting of the Hansard records changes over time. We use scripts written in R [@R2018] to convert the PDFs into daily text records.[^scriptsfootnote] Some error is introduced at this stage because many of the records are in a two-column format that need to be separated, and the PDF parsing is not always accurate, especially for older records. An example of the latter issue is that 'the' is often parsed as 'thc'. These errors are corrected when they can be identified, but as there are almost a billion words in the dataset, we are restricted to changes that can be made at scale.

[^scriptsfootnote]: Our code and data are available on request or via the GitHub repository for this paper: https://github.com/RohanAlexander/hansard. The scripts are primarily based on: the \texttt{PDFtools} R package of @Ooms2018pdftools; the \texttt{tidyverse} R package of @Wickham2017; the \texttt{tm} R package of @FeinererHornik2018; the \texttt{lubridate} R package of @GrolemundWickham2011; the \texttt{tidytext} R package of @SilgeRobinson2016; and the \texttt{stringi} R package of @Gagolewski2018. The functions of those packages are augmented by: the \texttt{furrr} R package of @VaughanDancho2018; and the \texttt{tictoc} R package of @Izrailev2014. The \texttt{hunspell} R package of @Ooms2017 is used to help find spelling issues; and the \texttt{quanteda} R package of @Benoit2018 is used to compound multiword expressions.

@@

[^datasourcefootnote]: Our dataset is available at: https://github.com/RohanAlexander/hansard. While we are making our data public in an attempt to help other researchers, we cleaned the dataset toward the requirements of this paper. If you have different requirements or have suggestions for how we can improve the dataset for other uses please contact us.

@@

We first use a Correlated Topic Model (CTM) [@BleiLafferty2007] to obtain estimated topic distributions over time, and then further group these topics to match those of the Comparative Agendas Project. We consider these topic distributions as inputs that can be analysed by another model. Thus, the second modelling step involves using a Bayesian hierarchical Dirichlet model to analyse changes in the topic distributions (obtained from the first step) in relation to events of interest.

@@

Other work such as @BaumgartnerJones1993 and @DowdingHindmoor2010 addressed this problem by creating a standardised codebook of policy categories and sub-categories and then manually assigning text to topics as appropriate. This approach ensures the categorisation is reasonable but as it is a manual process the size of the text that can be categorised is limited. @AshMorelliOsnabru2018 combined the best of both in the context of the New Zealand Hansard, using a supervised machine learning approach that would likely be beneficial in the Australian Hansard case as well.

In order to effectively categorise the topics of the entire Hansard, we use topic modelling, a statistical technique which aims to extract the underlying or latent 'topics' from a collection of texts. In particular, we use a method which is similar to Latent Dirichlet Allocation (LDA), first developed by @Blei2003latent. One advantage of using the Dirichlet distribution is that it formalises the idea that there is a trade-off in what is talked about in parliament given the limited time. A multinomial approach would not necessarily have this property.

@@

As such, we use the Correlated Topic Model (CTM) of @BleiLafferty2007 to obtain topic distributions. The CTM is a modification of LDA that allows for correlations between topics. More detail about the CTM is in Appendix \@ref(correlatedtopicmodelsection). There are various methods to help determine the appropriate number of topics to specify, and we further discuss the implications given our setting later in the paper. However the main implication is that the optimal number of topics was too large to be tractable by our analysis model, and too large to be easily visualised or examined. For this reason we manually group the topics identified by the CTM into those of the Comparative Agendas Project. It is these proportions that we use as an input to our analysis model.

@@

The output of interest from the topic modelling stage, is the proportion of each topic appearing in each document. The aim of the analysis stage of the modelling process is to analyse how the distribution of topics changes in relation to different types of events. But with many topics for each of the roughly 14,680 chamber-sitting-days the data are still too noisy to easily visualise changes around events.

@@

Define $\theta_{cdp}$ to be the proportion of topic $p$ on day $d$ in chamber $c$ (where chamber refers to either the House of Representatives or the Senate). Note that the $\theta_{c,d,1:P}$ for $p = 1,2,\dots, P = 19$ are equal to the estimated values of $\theta_{cd}$ from the CTM after we group them into the Comparative Agendas Project topics. We assume that the majority of variation in topics by day is across sitting periods, $s$, where a sitting period is defined as any group of days that are less than one week apart. Using this definition, there are 822 sitting periods over the period 1901 to 2018 inclusive. Appendix \@ref(hansardsummarystatistics) contains more information about the sitting patterns over the course of the year, which have changed considerably since Federation.

@@

The goal of the model is to relate these proportions to the current prime minister $g$ and election period $e$, assuming that an 'election effect' would be at its peak in the days close to the election, then decay over the period. The mean parameters $\mu_{c,s,p}$ are modelled on the log scale as:
$$
\log \mu_{c,s,p} = \alpha_{g[c,s],p} + \beta_{e[c,s],p} \cdot (N_{s[e]} - s) + \delta_{c,s,p}
$$
where: $\alpha_{g[c,s],p}$ is the mean effect for prime ministers $g$ (which covers sitting period $s$ in chamber $c$) and topic $p$; $\beta_{e[c,s],d,p}$ is the effect of election $e$ (which occurs in sitting period $s$, in chamber $c$) for topic $p$; $N_s$ is the total number of sitting periods in election period $e$; and $\delta_{c,s,p}$ is a random, or levels, effect for each chamber, sitting period and topic.

@@

Firstly, we describe the results of the CTM approach, which defined 80 topics over the period 1901 to 2018, and which we then further reduced to match the Comparative Agendas Project. We then describe the results of the Bayesian analysis model, which identified prime ministers, elections and other events that were associated with a change in the topics discussed.

@@

LDA output defines a topic as a distribution of probabilities over words. In Table \@ref(tab:topwordscomp) we display the ten words with the highest association for each of the 80 topics, but the topics are defined by probability distributions over all words. LDA does not apply labels to each topic, or collection of words, instead this has to be done by inspection. The topics cover areas such as budgets, demography, transport and infrastructure, war and conflict, health, education, agriculture, and trade. Similar to when topic models are run using the parliamentary text records of other countries, there are also some topics that seem to be about procedural or day-to-day matters, such as Topics 7 or 9. As expected, some topics seem to somewhat overlap with their content: for instance, Topics 4, 26, 28, 30 and 66 all relate to war and conflict.

@@

In the final column of Table \@ref(tab:topwordscomp) we summarise the results of our 80-topic model using the categories of the Comparative Agendas Project (CAP), as summarised by @Bevan2017.\footnote{We use Version 1.0 of the Master Codebook, as at 31 July 2014 available at \url{http://sbevan.com/cap-master-codebook.html}.} We do this firstly, to assist other researchers in understanding our results, by trying to relate them to a more well-known approach. We also do this because it allows us to further reduce our 80-topic model to only 21 topics.\footnote{Although the CAP Codebook topic numbering goes to 23, there are only 21 topics because there is no topic 11 or topic 22. Our 80 topics did not relate to CAP topics 21 or 23, meaning that we have 19 CAP-based categories.} It is these 19 CAP groupings to which we apply our analysis model.

@@

We are interested in considering the effect of various political and other events on what is talked about in the Australian Federal Parliament. As discussed in Section \@ref(analysismodelexplanation), the modelling process takes the topic proportions estimated by the CTM and reduced to the Comparative Agendas Project, and examines the association between these topic distributions and outside events. That is, the topic groupings are inputs for our analysis model.

There are several outputs of interest from this modelling stage. For example, the model provides estimates of topic prevalence by each sitting period. This nicely illustrates how the topics change over time, as the daily estimates tend to be quite variable, but using periods defined by prime ministers or elections tend not to provide enough variability. For instance, examining topics that have to do with war and conflict illustrates discussions of Australia's military involvement (Figure \@ref(fig:wartopicsgraph)).

```{r wartopicsgraph, cache = TRUE, echo = FALSE, fig.cap = "Model estimates of topic prevalance by sitting period for those related to war and conflict", out.width = '100%'}
knitr::include_graphics("../../results-analysis_model/two_house_results/defense_topic_fitted_data.pdf")
```

@@

```{r mugov, cache = TRUE, echo = FALSE, fig.cap = "Level effects for prime ministers by selected topics.", out.width = '90%'}
knitr::include_graphics("../../results-analysis_model/two_house_results/mu_gov_big_topics")
```

```{r muelection, cache = TRUE, echo = FALSE, fig.cap = "Level effects for election periods by selected topics", out.width = '90%'}
knitr::include_graphics("../../results-analysis_model/two_house_results/mu_election_big_topics.pdf")
```

@@

Table \@ref(tab:keyeventsresultstable) summarises the days where parliamentary discussion was significantly different from the rest prevailing in that week.

```{r keyeventsresultstable, echo = FALSE, results = 'asis'}

key_events_results_cols %>%
  kable(booktabs = T, caption = "Days that were significantly different to other days in their sitting period") %>%
  kable_styling(font_size = 10) %>%
  footnote(
    general = "These days were significantly different to others in their sitting period after taking election and prime minister effects into consideration.",
    footnote_as_chunk = T,
    threeparttable = TRUE)

@@

Of the 36 prime ministers over this period, we find that 17 of them were significantly different to the prime minister that preceded them, after we account for and remove some that had especially short terms. The three earliest of them occur in the first decade after Federation, and to a certain extent this may be due to the variety of the issues that had to be addressed in those initial years. Joseph Cook's time as prime minister is found to be significantly different to that of his predecessor, Andrew Fisher, although it may be that this is due to World War I.

@@

Menzies was succeeded  Holt in January 1966. This is an example where there was a change in prime minister without an election, as the next election only happened in November 1966. We find that Holt is different to Menzies.

<!-- In Figures \@ref(fig:muelectionall) we compare the topics during Menzies' final term with the topics of Holt. -->

<!-- ```{r muelectionall, cache = TRUE, echo = FALSE, fig.cap = "Differences between various elections", out.width = '100%'} -->
<!-- knitr::include_graphics("../../figures/mu_election-differences.pdf") -->
<!-- ``` -->

@@

One advantage of our analysis model compared with using the STM approach is that we can create a measure that is equivalent to testing for outliers in a model where the underlying variables were not latent. The results of this reduction in supervision are promising, but suggest the specifics of our process may need further refinement. For instance, our approach appropriately identifies the sitting day that first follows 11 September 2001. But there are many dates that we would have expected to be identified that were not, and similarly some of the dates that were identified are surprising. When we examine this we find that some of them are associated with significant legislation. Although our results are not overly over-weighted in the first half of our sample, there is a substantial gap between 1965 and 1991. Table \@ref(tab:keyeventsresultstable) highlights further work is needed to improve this approach. For instance, it may be that our approach is not appropriately considering step-changes or it may be that our identification of sitting periods is not appropriate for the entire period, given the changed sitting pattern that we identify.
